Distributional Trees and Forests for Circular Data

### 2019-12-09 pkg/circtree/inst/wind_paper/talks/CMStatistics2019/abstract/abstract_lang_cmstatistics2019.tex
Initial commit for CMStatistics2019 (abstract, slides, figs)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2923 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 4facf55885e7815d09efa4f5955432718f95ad6d

\author{Moritz N. Lang\\University of Innsbruck \And
        Lisa Schlosser\\University of Innsbruck \And
        Torsten Hothorn\\University of Zurich \And
        Georg J. Mayr\\University of Innsbruck \And
        Reto Stauffer\\University of Innsbruck \And
        Achim Zeileis\\University of Innsbruck}

\Abstract{
Circular data can be found in a variety of applications and subject areas,
e.g., hourly crime rate in the social-economics, animal movement direction or
gene-structure in biology, and wind direction as one of the most important
weather variables in meteorology. For probabilistic modeling of circular data
the von Mises distribution is widely used. While most of the already existing
approaches are built on additive regression models, we propose an adaption of
regression trees to circular data by employing distributional trees. In
comparison to the more commonly-used additive models, the resulting
distributional trees are easy to interpret, can detect non-additive effects,
and select covariates and their interactions automatically. In addition, as
natural extension ensembles or forests of such circular trees are introduced
that can further improve the forecasts by regularizing and stabilizing the
model. For illustration, short-term probabilistic wind direction forecasts at
airports are obtained in order to direct airplanes to a a safe landing. To
benchmark the performance of distributional trees and forests, their predictive
skill is compared to a persistence model, a climatology model and to a additive
regression model, employing the continuous ranked probability score. Model code
for circular distributional trees and forest is available in the R package "disttree".
}

\Keywords{}

\Plainkeywords{}

>>>>

\author{Moritz N. Lang\\University of Innsbruck 
\And Lisa Schlosser\\University of Innsbruck 
\And Torsten Hothorn\\University of Zurich 
\AND Georg J. Mayr\\University of Innsbruck 
\And Reto Stauffer\\University of Innsbruck 
\And Achim Zeileis\\University of Innsbruck}

\Plainauthor{Moritz N. Lang}

\Abstract{
Circular data can be found in a variety of applications and subject areas,
e.g., hourly crime rates in the social-economics, animal movement direction in
ecology, and wind direction as one of the most important weather variables in
meteorology. For probabilistic modeling of circular data the von Mises
distribution is widely used. While most existing approaches are built on
additive regression models, we propose an adaption of regression trees to
circular data by employing distributional trees. In comparison to the more
commonly-used additive models, the resulting distributional trees are easy to
interpret, can detect non-additive effects, and automatically select covariates
and their interactions. In addition, as a natural extension, ensembles or
forests of such circular trees are introduced that can further improve the
forecasts by regularizing and stabilizing the model. For illustration,
short-term probabilistic wind direction nowcasts at different airports are
obtained in order to direct airplanes to a safe landing. The predictive skill
of the novel approaches is benchmarked with an additive regression model plus a
persistency and a climatology model, employing the circular continuous ranked
probability score. The proposed methods for circular distributional trees and
forests are available in the R package ``disttree'' from R-Forge.
}

\Keywords{Distributional Trees; Distributional Forest; Circular Response; Von Mises Distribution}
